SIR Model With Testing Flows and Isolation

### 2020-11-05 codes/ali_test_sir.Rmd
back to initial sigma

### Assumptions
JD: Why are we assuming structurally that the test is perfectly specific? I guess it's better to put all of the boxes in and just sometimes assume that that $P_S=0$ 
- Assume there are 3 groups of individuals: $S$, $I$ and $R$.
- Each group splits into 4 subgroups to incorporate the testing mechanism, so overall the model framework has 12 compartments as follows. $S_u,$ $S_n$, $S_p$, $S_t$, $I_u$, $I_n$, $I_p$, $I_t$, $R_u$, $R_n$, $R_p$, $R_t$.
- Further, assume that (1) a perfectly specific test, i.e. no false positives, then $S_p$ and $S_t$ are always zero (2) there are two 'integrator' or 'accumulator' compartments, $N$ and $P$, to collect cumulative reported negative or positive tests. 
- There is a single FoI, $\Lambda$, all groups being exposed to all others. That is, where a susceptible individual move to does not depend on who infected it.

@@

- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively. We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.  It has units of $1/time$.
- $\gamma$: recovery rate ($1/time$).
- $\rho$: per capita testing intensity ($1/time$).
- $\tau$: the maximum rate of testing the whole untested population ($1/time$).
- $W$: weighted number of people available for tests, defined as $W = W_s S_u + W_i I_u + W_r R_u$.
- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\tau \rho N_0}{\tau W + \rho N_0}$.
- $F_z$: Weighted testing rate defined as $F_z=\sigma W_z$. That is, $F_s = \sigma W_s$, $F_i=\sigma W_i$ and $F_r = \sigma W_r$.

@@

- The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$. The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_z) F(W_z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$. Where, $P_z$ is the probability of positive test. Thus, $P_i$ represents the test accuracy. We assume $P_s=0$.

- On testing and scaling; at the simplest $F_z=W_z$.

- JD: If this $\tau$ is fast (e.g., 1d), it should have minimal effects on the system when we're in the usual domain. Since $\tau \gg \rho$, this generally collapses to the original form. When $W$ is super-small, however, it collapses instead to $\tau$.

@@

- Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
$\sigma = \frac{\rho N_0}{W}$ was used initially. The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$). This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_s=t N_0/S_u$. Thus, eq(1) of the model will be $d S_u/dt = - t N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.

### Model
The model is
\begin{align}
\label{mod:1}
1)\ d S_u/dt &= -\Lambda S_u - F_s S_u + \omega S_n \\
2)\ d S_n/dt &= -\Lambda S_n + F_s S_u - \omega S_n \\
3)\ d I_u/dt &= \Lambda S_u - F_i I_u + \omega I_n  - \gamma I_u  \\
4)\ d I_n/dt &= \Lambda S_n + (1-P_i) F_i I_u - \omega I_n -\gamma I_n \\
5)\ d I_p/dt &= P_i F_i I_u - \omega I_p -\gamma I_p \\
6)\ d I_t/dt &= \omega I_p - \gamma I_t  \\
7)\ d R_u/dt &= \gamma I_u - F_r R_u + \omega R_n \\
8)\ d R_n/dt &= \gamma I_n + (1-P_r) F_r R_u - \omega R_n  \\
9)\ d R_p/dt &= \gamma I_p + P_r F_r R_u  - \omega R_p  \\
10)\ d R_t/dt&= \gamma I_t + \omega R_p  \\
11)\ dN/dt &= \omega (S_n + I_n + R_n)   \\
12)\ dP/dt &= \omega(I_p + R_p).
\end{align}

@@

```{r, model formulation}
sir.model <- function(time,state,params){
    unpack(as.list(c(state,params)))
    ## Force of Infection  
    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_t*I_t)/N0
    ## scaling the weights
    sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
    F_s <- sigma*W_s
    F_i <- sigma*W_i
    F_r <- sigma*W_r

    dS_u.dt <- -Lambda*S_u - (1-P_s) * F_s * S_u + omega * S_n
    dS_n.dt <- -Lambda*S_n + (1-P_s) * F_s * S_u - omega * S_n 
    dI_u.dt <-  Lambda*S_u + omega * I_n - F_i * I_u - gamma * I_u
    dI_n.dt <-  Lambda*S_n + (1-P_i) * F_i * I_u  - omega * I_n - gamma * I_n
    dI_p.dt <-  P_i * F_i * I_u - omega * I_p - gamma * I_p 
    dI_t.dt <-  omega * I_p - gamma * I_t 
    dR_u.dt <- gamma * I_u + omega * R_n - F_r * R_u  
    dR_n.dt <- gamma * I_n + (1-P_r) * F_r * R_u - omega * R_n
    dR_p.dt <- gamma * I_p + P_r * F_r * R_u - omega * R_p
    dR_t.dt <- gamma * I_t + omega * R_p
    dN.dt <- omega * (S_n + I_n + R_n)
    dP.dt <- omega *(I_p + R_p)

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=0.8,gamma=1/3, omega=0.25, rho=0.01, 
             W_s=.01, W_i=1, W_r=1, #weights
             P_s=0, P_i=1, P_r=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2, #eta_w/eta_t
            tau=1
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
rootfun <- function (time, state,params) { 
  unpack(as.list(c(state,params)))
  return(I_u - 1) }
```

```{r DFE, message=FALSE}
# DFE; numerical approach
library(rootSolve)
DFE <- function(S,params){
  S_u <- S[1]
  S_n <- S[2]
  I_u <-0
  R_u<-0
  unpack(as.list(params))
  sigma <- tau*rho*N0/(tau*(W_s*S_u+W_i*I_u+W_r*R_u)+rho*N0)
  F_s <- sigma*W_s

 return(c(eq1=S_u+S_n-N0,
   eq2=-F_s*S_u+omega*S_n))
}

@@

# DFE; analytic approach

Su_dfe <-function(params){
  unpack(as.list(params))
  a <- rho
  b <- -rho+(1-rho/omega)*tau*W_s
  c <- -tau*W_s
  d <- b^2-4*a*c

  if(rho){return(N0*2*a/(-b+sqrt(d)))}
    else {return(N0)}
}

Sn_dfe<- params[["N0"]]-Su_dfe(params)

state_dfe <- c(S_u=Su_dfe(params), S_n=Sn_dfe,
                I_u=0,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)

all.equal(c(state_dfe[["S_u"]],state_dfe[["S_n"]]),roots_dfe$root)
```

Test gradient
```{r}
g1 <- sir.model(time=0,state=state_init, params=params)
all.equal(sum(g1[[1]]),0)
## debug(sir.model)
## g1 <- sir.model(time=0,state=state_init, params=params)
```

@@

The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is given by the solution of the following system:
$S_u+S_n=N0$ and $F_s S_u-\omega S_n=0$.
Let $\tilde{S_u}=N_0/S_u$, the positive solution of the quadratic $$a \tilde{S_u}+b \tilde{S_u}^2+c=0,$$
where $a=\rho$, $b= -\rho+(1-\rho/\omega) \tau W_s$ and $c=-\tau W_s$. Let
  $d = b^2-4ac$, the DFE of the model is

$$S_u^* =N_0 2 a/(-b+\sqrt d), \\ S_n^*=N_0-S_u^* \\ I_j=R_j=0 \ \forall j$$

@@

$$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t P_i) F_i$,
$B=\big(\omega+(F_i+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega P_i F_i }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_i(\gamma+\omega P_i))}$.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_i+\gamma&-\omega&0&0\\
-(1-P_i)F_i&\omega+\gamma&0&0\\
-P_i F_i&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

@@

The Eigenvalues of $G_{11}$, are 0 and $R_0$.
```{r R0}
F_i <- function(params){
   # Returns F_i at DFE 
  unpack(as.list(params))
  Su_dfe<- Su_dfe(params)
  Sn_dfe<- N0-Su_dfe
  sigma <- tau*rho*N0/(tau*(W_s*Su_dfe)+rho*N0) #at DFE 
  return(sigma*W_i)
}

 # just a test
  F_i(params=update(params,rho=0))

R0<-function(params){
    unpack(as.list(params))
    Su_dfe<- Su_dfe(params)
    Sn_dfe<- N0-Su_dfe
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    Fi <- F_i(params) #at DFE

    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*P_i*Fi
    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*P_i*Fi)/(omega+gamma)

    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*P_i)))
    return((A*Su_dfe+B*Sn_dfe)*C)
}

# R0 expression with eta_w/eta_t factored
R02<-function(params){
    unpack(as.list(params))
    Su_dfe<- Su_dfe(params)
    Sn_dfe<- N0-Su_dfe
    F_i <-F_i(params)  #at DFE

    A2 <- gamma*(P_i*F_i*omega*Sn_dfe+(omega+gamma)*(gamma*Sn_dfe+F_i*N0))
    B2 <- P_i*F_i*omega*(gamma*Su_dfe+omega*N0)
    D <- gamma*(omega+gamma)*(gamma*Su_dfe+omega*N0)
    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_i*(gamma+omega*P_i)))*(omega+gamma))
    # s <- eta_w/eta_t

    # return((eta_w*A2+eta_t*B2+D)*C2)
    return((eta_t*(s*A2+B2)+D)*C2)
}

@@

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.001,1 , length.out=n_in) #note omega must be > rho, was 0.1

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]],tau=params[["tau"]])

# Need to be corrected: in F_i, state_dfe should be changed according to the df1!
df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
                      Fi= apply(df1,1,function(params_in)F_i(params=params_in)))

# which(df2[]<0)
```

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
 # beta  <- 0.5
 # eta_w <- seq(0,1, length.out=5) 
 # eta_t <- seq(0.001,0.5, length.out=n_out) #0.001 
 # rho     <- seq(0,0.01, length.out=n_in)
 # omega <- seq(0.001,1 , length.out=n_in) #note omega must be > 0, was 0.1
 s <- seq(1,2,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_s=params[["W_s"]],W_i=params[["W_i"]],W_r=params[["W_r"]],
                    P_s=params[["P_s"]],P_i=params[["P_i"]],P_r=params[["P_r"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s,tau=params[["tau"]])

df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
                      Fi= apply(df11,1,function(params_in)F_i(params=params_in)),
                       Su=apply(df11,1,function(params_in)Su_dfe(params = params_in)))
```

@@

Let call the $F_i$ the critical testing rate in the above plot.

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_s} \Big(\gamma(\eta_w-1)(\gamma W_s+\omega W_i) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_s\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_i-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_s$ and $c=(\eta_w-1)\gamma^2 W_s$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

4. Taylor expansion gives $R_0|_{(\rho=0,W_s=1)} = \beta/\gamma$.

5. Taylor expansion gives $R_0|_{(\rho=0,W_s=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_i+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_s+\tau\rho(\gamma+\omega P_i)W_i+ \gamma \rho(\gamma+\omega)}$.

>>>>

### Assumptions
JD: Why are we assuming structurally that the test is perfectly specific? I guess it's better to put all of the boxes in and just sometimes assume that that $p_S=0$ 
- Assume there are 3 groups of individuals: $S$, $I$ and $R$.
- Each group splits into 4 subgroups to incorporate the testing mechanism, so overall the model framework has 12 compartments as follows. $S_u,$ $S_n$, $S_p$, $S_t$, $I_u$, $I_n$, $I_p$, $I_t$, $R_u$, $R_n$, $R_p$, $R_t$.
- Further, assume that (1) a perfectly specific test, i.e. no false positives, then $S_p$ and $S_t$ are always zero (2) there are two 'integrator' or 'accumulator' compartments, $N$ and $P$, to collect cumulative reported negative or positive tests. 
- There is a single FoI, $\Lambda$, all groups being exposed to all others. That is, where a susceptible individual move to does not depend on who infected it.

@@

- $\Lambda$: force of infection defined as $$\Lambda=\beta \frac{(I_u+\eta_w I_n+\eta_w I_p+ \eta_t I_t)}{N_0},$$ where $\beta$ is transmission rate, $\eta_w$ and $\eta_t$ are the isolation parameters for awaiting and reported individuals, respectively. We assume that $\eta_t<\eta_w$, i.e., the awaiting individuals for test results have a higher transmission probability than the reported individuals.
- $\omega$: the rate of *onward flow* from the awaiting positive compartment, $p$, to reported/tested compartment, $\rho$, or from awaiting negative compartment, $n$, back to $u$.  It has units of $1/time$.
- $\gamma$: recovery rate ($1/time$).
- $\rho$: per capita testing intensity across the whole population ($1/time$).
- $W$: weighted number of people available for tests, defined as $W = W_S S_u + W_I I_u + W_R R_u$.
- $\sigma$: scaling parameter for testing defined as $\sigma = \frac{\rho N_0}{W}$.
- $F_Z$: Weighted testing rate defined as $F_Z=\sigma W_Z$. That is, $F_S = \sigma W_S$, $F_I=\sigma W_I$ and $F_R = \sigma W_R$.

@@

- The complicated part is the flow rate from $Z_u$ to $Z_n$, $Z_p$. The (total, not per capita) flow from $Z_u$ to $Z_n$ is $\rho (1-P_Z) F(W_Z) Z_u$, from $Z_u$ to $Z_p$ is $\rho P_Z F(W_Z) Z_u$. Where, $P_Z$ is the probability of positive test. Thus, $p_I$ represents the test accuracy. We assume $p_S=0$.

- On testing and scaling; at the simplest $F_Z=W_Z$.

@@

- Initial approach and singularity issue (not sure if I call it structural or numerical singularity?); 
$\sigma = \frac{\rho N_0}{W}$ was used initially. The issue was that the population in $S$ compartments appeared to blow up when the DFE is achieved ( in the simulation when time, $t>200$). This is  once the only untested people are susceptibles, the FoI will become $\Lambda=0$, testing rate $F_S=\rho N_0/S_u$. Thus, eq(1) of the model will be $d S_u/dt = - \rho N_0 + \omega S_n$ which is no longer dependent on $S_u$ and a linear rate of leaving the $S_u$ compartment.

### Model
The model is
\begin{align}
\label{mod:1}
1)\ d S_u/dt &= -\Lambda S_u - F_S S_u + \omega S_n \\
2)\ d S_n/dt &= -\Lambda S_n + F_S S_u - \omega S_n \\
3)\ d I_u/dt &= \Lambda S_u - F_I I_u + \omega I_n  - \gamma I_u  \\
4)\ d I_n/dt &= \Lambda S_n + (1-p_I) F_I I_u - \omega I_n -\gamma I_n \\
5)\ d I_p/dt &= p_I F_I I_u - \omega I_p -\gamma I_p \\
6)\ d I_t/dt &= \omega I_p - \gamma I_t  \\
7)\ d R_u/dt &= \gamma I_u - F_R R_u + \omega R_n \\
8)\ d R_n/dt &= \gamma I_n + (1-p_R) F_R R_u - \omega R_n  \\
9)\ d R_p/dt &= \gamma I_p + p_R F_R R_u  - \omega R_p  \\
10)\ d R_t/dt&= \gamma I_t + \omega R_p  \\
11)\ dN/dt &= \omega (S_n + I_n + R_n)   \\
12)\ dP/dt &= \omega(I_p + R_p).
\end{align}

@@

```{r, model formulation}
sir.model <- function(time,state,params){
    unpack(as.list(c(state,params)))
    ## Force of Infection  
    Lambda <- beta * (I_u + eta_w*(I_n+I_p) + eta_t*I_t)/N0
    ## scaling the weights
    W <- W_S*S_u+W_I*I_u+W_R*R_u
    sigma <- rho*N0/W
    F_S <- sigma*W_S
    F_I <- sigma*W_I
    F_R <- sigma*W_R

    dS_u.dt <- -Lambda*S_u - (1-p_S) * F_S * S_u + omega * S_n
    dS_n.dt <- -Lambda*S_n + (1-p_S) * F_S * S_u - omega * S_n 
    dI_u.dt <-  Lambda*S_u + omega * I_n - F_I * I_u - gamma * I_u
    dI_n.dt <-  Lambda*S_n + (1-p_I) * F_I * I_u  - omega * I_n - gamma * I_n
    dI_p.dt <-  p_I * F_I * I_u - omega * I_p - gamma * I_p 
    dI_t.dt <-  omega * I_p - gamma * I_t 
    dR_u.dt <- gamma * I_u + omega * R_n - F_R * R_u  
    dR_n.dt <- gamma * I_n + (1-p_R) * F_R * R_u - omega * R_n
    dR_p.dt <- gamma * I_p + p_R * F_R * R_u - omega * R_p
    dR_t.dt <- gamma * I_t + omega * R_p
    dN.dt <- omega * (S_n + I_n + R_n)
    dP.dt <- omega *(I_p + R_p)

@@

```{r model inputs, message=FALSE}
params <- c(N0=1000000, beta=0.5,gamma=1/3, omega=0.25, rho=0.01, 
             W_S=1/3, W_I=1/3, W_R=1/3, #weights
             p_S=0, p_I=1, p_R=0.5, #prob of waiting for being tested positive
            eta_w=0.02, eta_t=0.01, #isolation parameter 
            s=2 #eta_w/eta_t
            )
class(params) <- "params_pansim"
state_init <- c(S_u=params[["N0"]], S_n=0,
                I_u=1,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
d <- seq(0,200,by=0.1)
# Conditional stop for desolver
# rootfun <- function (time, state,params) { 
#   unpack(as.list(c(state,params)))
#   return(I_u - 1) }
```

```{r DFE, message=FALSE}
# DFE; numerical approach
library(rootSolve)
DFE <- function(S,params){
  S_u <- S[1]
  S_n <- S[2]
  I_u <-0
  R_u <-0
  unpack(as.list(params))
  W <- W_S*S_u+W_I*I_u+W_R*R_u
  sigma <- rho*N0/W
  F_S <- sigma*W_S

 return(c(eq1=S_u+S_n-N0,
          eq2=-F_S*S_u+omega*S_n))
}

@@

# DFE; analytic approach
Sn_dfe <-function(params){
  unpack(as.list(params))
  return((rho*(1-p_S)*N0)/omega)
  }

Su_dfe<- params[["N0"]]-Sn_dfe(params)

state_dfe <- c(S_u=Su_dfe, S_n=Sn_dfe(params),
                I_u=0,I_n=0,I_p=0,I_t=0,
                R_u=0,R_n=0,R_p=0,R_t=0,
                N=0,P=0)
# check:
all.equal(c(state_dfe[["S_u"]],state_dfe[["S_n"]]),roots_dfe$root)
```

Test gradient
```{r}
g1 <- sir.model(time=0,state=state_init, params=params)
all.equal(sum(g1[[1]]),0)
```

@@

The Disease-Free Equilibrium, namely $E_1$, for the above SIR model is given by the solution of the following system:
$S_u+S_n=N0$ and $F_S S_u-\omega S_n=0$. The DFE is

$S_n^*= \frac{\rho}{\omega} N_0, S_u^*= N_0-S_n^*$, and $I_j=R_j=0 \ \forall j$.

@@

$$R_0= (A \times S_u^* + B \times S_n^*) \times C,$$ where
$A=\gamma(\omega+\gamma) + (\gamma \eta_w + \omega \eta_t p_I) F_I$,
$B=\big(\omega+(F_I+\gamma)\eta_w\big) \gamma+\frac{(\eta_w \gamma+ \eta_t\omega) \omega p_I F_I }{\omega+\gamma}$ and 
$C=\frac{\beta/\gamma}{N_0 (\gamma(\omega+\gamma)+F_I(\gamma+\omega p_I))}$.

\begin{equation}\label{FV}
F= \beta/N_0 \left[ \begin {array}{cccc} 
S_u&\eta_w\,S_u&\eta_w\,S_u&\eta_t\,S_u\\
S_n&\eta_w\,S_n&\eta_w\,S_n&\eta_t\,S_n\\ 
0&0&0&0\\
0&0&0&0
 \end {array} \right],
 V=
 \left[ \begin {array}{cccc}  
F_I+\gamma&-\omega&0&0\\
-(1-p_I)F_I&\omega+\gamma&0&0\\
-p_I F_I&0&\omega+\gamma&0\\
0&0&-\omega&\gamma
\end {array} \right],
\end{equation}

@@

The Eigenvalues of $G_{11}$, are 0 and $R_0$.
```{r R0}
F_I <- function(params){
   # Returns F_I at DFE 
  unpack(as.list(params))
  Sn<- Sn_dfe(params)
  Su<- N0-Sn
  sigma <- rho*N0/(W_S*Su) #at DFE 
  return(sigma*W_I)
}

 # just a test
  F_I(params=update(params,rho=0))

R0<-function(params){
    unpack(as.list(params))
    Sn <- Sn_dfe(params)
    Su <- N0-Sn
    # print(c(Su_dfe,Sn_dfe,Su_dfe+Sn_dfe))
    Fi <- F_I(params) #at DFE
    A <-gamma*(omega+gamma+eta_w*Fi) + omega*eta_t*p_I*Fi
    B <-(omega+eta_w*(Fi+gamma))*gamma+(eta_w*gamma+eta_t*omega)*(omega*p_I*Fi)/(omega+gamma)
    C <- beta/(N0*gamma*(gamma*(omega+gamma)+Fi*(gamma+omega*p_I)))
    return((A*Su+B*Sn)*C)
}

# R0 expression with eta_w/eta_t factored
R02<-function(params){
    unpack(as.list(params))
    Sn<- Sn_dfe(params)
    Su<- N0-Sn
    F_I <-F_I(params)  #at DFE

    A2 <- gamma*(p_I*F_I*omega*Sn+(omega+gamma)*(gamma*Sn+F_I*N0))
    B2 <- p_I*F_I*omega*(gamma*Su+omega*N0)
    D <- gamma*(omega+gamma)*(gamma*Su+omega*N0)
    C2 <- beta/((N0*gamma*(gamma*(omega+gamma)+F_I*(gamma+omega*p_I)))*(omega+gamma))
    # s <- eta_w/eta_t
    return((eta_t*(s*A2+B2)+D)*C2)
}

@@

### Contour plot of $R_0=1$ and interpretations approache 1
```{r contour_R0_1, message=FALSE,warning=FALSE }
n_out <- 5 ## facets (rows/cols)
n_in <- 41 ## grid N within facets
 # specify the ranges
 beta  <- 0.5  ## set beta high enough to allow R0>1 in worst case
 eta_w <- seq(0,1, length.out=n_out)
 eta_t <- seq(0,1, length.out=n_out) #0.001 to 0.5 
 rho     <- seq(0,0.01, length.out=n_in)
 omega <- seq(0.011,1 , length.out=n_in) #note omega must be > rho, was 0.1

 df1 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=eta_w,eta_t=eta_t,s=params[["s"]])

# Need to be corrected: in F_I, state_dfe should be changed according to the df1!
df2<- data.frame(df1,R0=apply(df1,1,function(params_in)R0(params=params_in)),
                      Fi= apply(df1,1,function(params_in)F_I(params=params_in)))

which(df2[]<0)
```

@@

```{r contour_R0_2, message=FALSE,warning=FALSE }
 # specify the ranges
eta_t <- seq(0.001,0.5, length.out=n_out)  
s <- seq(1,2,length.out=n_out)

 df11 <- expand.grid(N0=params[["N0"]],beta=beta,gamma=params[["gamma"]],omega=omega,rho=rho,
                    W_S=params[["W_S"]],W_I=params[["W_I"]],W_R=params[["W_R"]],
                    p_S=params[["p_S"]],p_I=params[["p_I"]],p_R=params[["p_R"]],
                    eta_w=params[["eta_w"]],eta_t=eta_t,s=s)

df22<- data.frame(df11,R0=apply(df11,1,function(params_in)R02(params=params_in)),
                      Fi= apply(df11,1,function(params_in)F_I(params=params_in)),
                       Sn=apply(df11,1,function(params_in)Sn_dfe(params = params_in)))
```

@@

Let call the $F_I$ the critical testing rate in the above plot.

@@

1. $\partial{R_0}/\partial{\eta_t} \geq 0$ and $\partial{R_0}/\partial{\eta_w} \geq 0$.
2. $\partial{R_0}/\partial{\rho} \leq 0$; this is concluded from the Taylor expansion of $R_0$ around $\rho=0$ which is
$R_0 \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 W_S} \Big(\gamma(\eta_w-1)(\gamma W_S+\omega W_I) + (\eta_t -1)P_iW_i \omega^2 \Big) + \mathcal{O}(\rho^2)$.
3. Based on the above Taylor expansion around $\rho=0$,
$$\partial{R_0}/\partial{\omega}=  \frac{-\beta \rho}{\gamma W_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c), $$
where $a=(\eta_w-1)W_I-(\eta_t-1)P_iW_i$, $b=2(\eta_w-1)\gamma W_S$ and $c=(\eta_w-1)\gamma^2 W_S$.
Given that $\eta_t\leq \eta_w$, it is inferred that $a\geq 0$, $b\leq 0$ and $c \leq 0$.
Let $\omega_1<0$ and $\omega_2>0$ be the roots of the quadratic expression in $\partial{R_0}/\partial{\omega}$. Thus, $\partial{R_0}/\partial{\omega}>0$ for $0<\omega<\omega_2$ and  $\partial{R_0}/\partial{\omega}<0$ for $\omega>\omega_2$.

4. Taylor expansion gives $R_0|_{(\rho=0,W_S=1)} = \beta/\gamma$.

5. Taylor expansion gives $R_0|_{(\rho=0,W_S=0)} = \beta/\gamma \frac{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma\eta_w+\omega\eta_tP_i)W_I+ \gamma \rho(\gamma+\omega)}{\gamma \tau(\omega+\gamma)W_S+\tau\rho(\gamma+\omega p_I)W_I+ \gamma \rho(\gamma+\omega)}$.
